---
title: "Car Price Prediction"
author: "Jiachen_Liu and Thinh_Nguyen"
date: "4/15/2021"
output: 
  html_document: 
    fig_width: 10
    fig_height: 9
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
```
## EDA 
```{r}
car = read.csv('raw_data.csv')
```

```{r}
head(car)
```


```{r}
## choose subset of bmw data
bmw = car[which( car$brand == 'bmw' ),]
bmw = subset(bmw,select = -1 )
head(bmw)
```


```{r}
## summary of bmw data 
summary(bmw)
```
```{r}
## check missing value for each column
library(mice)
library(VIM)
md.pattern(bmw)
aggr(bmw, prop=FALSE, numbers=TRUE)
```
```{r}
colnames(bmw)
```
## Exploary Data Analysis 

```{r}
# box plot for the bmw year
ggplot(data = bmw, aes( x = year, y = price, group =year ) ) + geom_boxplot(color = "blue",alpha = 0.1 ) + ggtitle("Box plot of bmw year")+ theme_minimal()
```

```{r}
# box plot of bmw  price
ggplot(data = bmw, aes(x = transmission, y = price )) + geom_boxplot( aes(color = transmission), alpha = 0.1) + ggtitle("Box plot of bmw price")+ theme_minimal()
```


```{r}
# boxplot for car model
ggplot(data = bmw, aes( y = price, x = model, group = model ) ) + geom_boxplot( aes(color = model), alpha =0.8 ) + ggtitle("Box plot of car mileage in different model")

```


```{r}
# scatter plot of millage and price

ggplot(data = bmw, aes(x = mileage , y = price)) + geom_point(color ="red", alpha= 0.05) + ggtitle("The relationship between millage and price in different transmission")+ facet_wrap(~ transmission)

```


```{r}
#scatter plot for car mpg and price
ggplot(data = bmw, aes(x = engineSize , y = price , group = engineSize)) + geom_boxplot(color ="blue", alpha= 0.05) + ggtitle("The relationship between engine size and price")
```


```{r}
ggplot(data = bmw[which(bmw$year>2014),], aes(x = mileage , y = price)) + geom_point(color ="red", alpha= 0.05) + ggtitle("The relationship between millage and price in different transmission")+ facet_wrap(~ year  )

```

```{r}
# bar plot of "Total Number of each transmission"
ggplot( data = bmw , aes(x = transmission ), color = "red" ) + geom_bar(aes(color = transmission)) + ggtitle( "Total Number of each transmission")+ theme_minimal()

```





## Feature Engineering

### encoding categrical data
```{r}
# Feature Engineering 
library(caret)
library(psych)
library(fastDummies)
```

```{r}
bmw_data  = dummy_cols(bmw, select_columns =c('year','model','transmission','fuelType','engineSize' ) )
head(bmw_data)
```

```{r}
bmw_df = subset(bmw_data ,select =c(-year,-model, -transmission, - fuelType, -brand , -engineSize) )
head(bmw_df)
```


### Normalize data and prepare for PCA
```{r}
bmw_scale = as.data.frame(scale(bmw_df[  , -2]))
head(bmw_scale)
```

### Apply PCA on scale data set and choose apporiate number of components
```{r}
fa.parallel(bmw_scale,fa = "pc",n.iter = 100,show.legend = TRUE )

```

### Choose components number is 31 and create pca data frame for modeling part
```{r}
# From the result from above , we choose components number = 31 
pc = principal(bmw_scale, nfactors = 31 , rotate = "none", scores = TRUE )
```
```{r}
pca_df = cbind(bmw_df$price , as.data.frame(pc$scores))%>%
rename(price = "bmw_df$price")
head(pca_df)
```

### Split dataset as train and test data set
```{r}
# split train and test data , which 80% train data and 20% test data
library(caTools)
pca_split_data = sample.split(pca_df,SplitRatio = 0.8)
pca_train_data = subset(pca_df,pca_split_data==TRUE)
pca_test_data = subset( pca_df,pca_split_data ==FALSE)

```


## implementing model 

### linear regression with PCA
```{r}
## linear regression 
linear = lm(price ~. , data = pca_train_data)
summary(linear)
plot(linear)
```


### Lasso Regression with PCA
```{r}
## Lasso Regression
lasso_pca <- train(price ~ . ,
                data = pca_train_data, 
                method = "lasso")
lasso_pca
plot(lasso_pca)
```




### Ridge Regression with PCA
```{r}
## ridge regression
ridge_pca <- train(price ~ . ,
                data = pca_train_data, 
                method = "ridge")
ridge_pca
plot(ridge_pca)
```


### Elasticnet with PCA
```{r}
## Elasticnet
enet_pca <- train(price ~ . ,
                data = pca_train_data, 
                method = "enet")
enet_pca
plot(enet_pca)
```



### KNN regression with PCA
```{r}
## KNN regression
knn_pca <- train(price ~ . ,
                data = pca_train_data, 
                method = "knn")
knn_pca
plot(knn_pca)
```


### XGBoost with PCA
```{r}
## xgboost
xgboost_pca<- train(price ~ . ,
                data = pca_train_data, 
                method = "xgbTree")
xgboost_pca
plot(xgboost_pca)
```


## Model Evaluation
### Linear Regression on test set
```{r}
## RMSE, R2 and MAE in test data set of Linear Regression 
linear_pca_pred <- predict(linear, pca_test_data)
as.table(postResample(pred = linear_pca_pred, obs = pca_test_data$price))
ggplot() + geom_point(aes(x = pca_test_data$price , y = linear_pca_pred), color = "red" ,alpha =0.2) + xlab("True_Price") + ylab("Pred_Price") + ggtitle("The scatter plot of true price and predict price(Linear Regression)")+geom_abline (slope= 1, color = "blue")
```


### Lasso Regression on test set
```{r}
## RMSE, R2 and MAE in test data set of Linear Regression
lasso_pca_pred <- predict(lasso_pca, pca_test_data)
as.table(postResample(pred = lasso_pca_pred, obs = pca_test_data$price))
ggplot() + geom_point(aes(x = pca_test_data$price , y = lasso_pca_pred), color = "red" ,alpha =0.2) + xlab("True_Price") + ylab("Pred_Price") + ggtitle("The scatter plot of true price and predict price(Lasso Regression)")+geom_abline (slope= 1, color = "blue")

```

### Ridge Regression on test set
```{r}
## RMSE, R2 and MAE in test data set of Ridge Regression
ridge_pca_pred <- predict(ridge_pca, pca_test_data)
as.table(postResample(pred = ridge_pca_pred, obs = pca_test_data$price))
ggplot() + geom_point(aes(x = pca_test_data$price , y = ridge_pca_pred), color = "red" ,alpha =0.2) + xlab("True_Price") + ylab("Pred_Price") + ggtitle("The scatter plot of true price and predict price(Ridge Regression)")+geom_abline (slope= 1, color = "blue")
```

### Elasticnet on test set 
```{r}
## RMSE, R2 and MAE in test data set of Elastic net
enet_pca_pred <- predict(enet_pca, pca_test_data)
as.table(postResample(pred = enet_pca_pred, obs = pca_test_data$price))
ggplot() + geom_point(aes(x = pca_test_data$price , y = enet_pca_pred), color = "red" ,alpha =0.2) + xlab("True_Price") + ylab("Pred_Price") + ggtitle("The scatter plot of true price and predict price(Elastic net Regression)")+geom_abline (slope= 1, color = "blue")

```

### KNN Regression on test set 
```{r}
## RMSE, R2 and MAE in test data set 
knn_pca_pred <- predict(knn_pca, pca_test_data)
as.table(postResample(pred = knn_pca_pred, obs = pca_test_data$price))
ggplot() + geom_point(aes(x = pca_test_data$price , y = knn_pca_pred), color = "red" ,alpha =0.2) + xlab("True_Price") + ylab("Pred_Price") + ggtitle("The scatter plot of true price and predict price(KNN Regression)")+geom_abline (slope= 1, color = "blue")
```

### XGboost on test set 
```{r}
## xgboost evaluation 
xgboost_pca_pred <- predict(xgboost_pca, pca_test_data)
as.table(postResample(pred = xgboost_pca_pred, obs = pca_test_data$price))
ggplot() + geom_point(aes(x = pca_test_data$price , y = xgboost_pca_pred), color = "red" ,alpha =0.2) + xlab("True_Price") + ylab("Pred_Price") + ggtitle("The scatter plot of true price and predict price(XGBoost)")+geom_abline (slope= 1, color = "Blue")

```













